### 2016-06-18 paper/icfp16-cameraready/evaluation.tex
number of exams
add threats to validity

\begin{figure*}[ht]
\centering
\begin{minipage}{\linewidth}
\centering
\includegraphics[width=0.49\linewidth]{coverage.png}
% \end{minipage}
% \begin{minipage}{\linewidth}
\includegraphics[width=0.49\linewidth]{distrib.png}
\end{minipage}
% \vspace{-8ex}
\caption{Results of our coverage testing and the distribution of test
  outcomes. Our random search successfully finds witnesses for 79--85\% of
  the programs in under one second, improving slightly to 87\% in under
  10 seconds. In both datasets we detect actual type errors about 82\%
  of the time, unbound variables or constructors 3--4\% of the time, and
  diverging loops 2--3\% of the time. For the remaining 11--12\% of the
  programs we are unable to provide any useful feedback.  }
\label{fig:results-witness}
\end{figure*}

\paragraph{Results}
\label{sec:results-witness}
The results of our experiments are summarized in
Figure~\ref{fig:results-witness}.
%
In both datasets our tool was able to find a witness for 83\% of the
programs in under one second, \ie\ fast enough to be integrated as a
compile-time check. If we extend our tolerance to a 10 second timeout,
we hit a maximum of 87\% coverage.
%
Interestingly, while the vast majority of witnesses corresponded to a
type-error, as expected, 3--4\% triggered an unbound variable error (even
though \ocaml\ reported a type error) and 2--3\% triggered an infinite
recursion error.
%
For the remaining 12\% of programs we were unable to provide any useful
feedback as they either completed 1,000 tests successfully, or timed out
after one minute.
%
% XX programs were deemed safe and XX timed out even at 3,000 steps, \ie
% we could not provide any useful feedback for XX\% of the total programs.
%
While a more advanced search procedure, \eg\ dynamic-symbolic execution,
could likely trigger more of the type errors, our experiments show that
type errors are coarse enough (or that novice programs are \emph{simple}
enough) that these techniques are not necessary.

@@

\subsection{Measuring Witness Utility}
\label{sec:user-study}
Finally, to test the explanatory power of our jump-compressed traces, we
ran a user study (IRB \#XXXX) at the University of Virginia (UVA).
%
We included four problems in an exam in the Spring session of UVA's
undergraduate Programming Languages course (CS XXX).
%
The problems presented the students with ill-typed \ocaml\ programs and
asked them to
%
(1) \emph{explain} the type error, and
%
(2) \emph{fix} the type error.
%
For each problem the student would be given the ill-typed program and
either \ocaml's error message or \toolname's jump-compressed trace.
%
The four problems assigned to the students were the @sumList@,
@digitsOfInt@, and @wwhile@ programs from \S~\ref{sec:advantage-traces},
as well as the following @append@ program
%
\begin{ecode}
  let append x l =
    match x with
    | []   -> l
    | h::t -> h :: t :: l
\end{ecode}
%
which triggers an occurs-check error in the ``cons'' case.
%
We then instructed four annotators (one of whom is an author, the other
three are teaching assistants at UCSD) to classify the answers as
correct or incorrect.

@@

\paragraph{Results}
%
Figure~\ref{fig:results-user-study} summarizes a single annotator's
results, which show that students given \toolname's jump-compressed
trace were consistently more likely to correctly explain and fix the
type error than those given \ocaml's error message.
%
The measured kappa values were $\kappa = 0.72$ for the explanations and
$\kappa = 0.83$ for the fixes; while there is no formal notion for what
consititutes strong agreement~\cite{Krippendorff2012-wd}, kappa values
above $0.60$ are often called ``substantial''
agreement~\cite{Landis1977-ey}.
%
\begin{figure*}[ht]
\centering
\includegraphics[width=0.49\linewidth]{user-study-reason.png}
\includegraphics[width=0.49\linewidth]{user-study-fix.png}
\caption{A classification of students' explanations and fixes for type
  errors, given either \ocaml's error % message
  or \toolname's
  jump-compressed trace. The students given \toolname's jump-compressed
  trace consistently scored better than those given \ocaml's type
  error.}
\label{fig:results-user-study}
\end{figure*}
%

>>>>

\begin{figure*}[ht]
\centering
\begin{minipage}{\linewidth}
\centering
\includegraphics[width=0.49\linewidth]{coverage.png}
% \end{minipage}
% \begin{minipage}{\linewidth}
\includegraphics[width=0.49\linewidth]{distrib.png}
\end{minipage}
% \vspace{-8ex}
\caption{Results of our coverage testing and the distribution of test
  outcomes. Our random search successfully finds witnesses for 79--85\% of
  the programs in under one second, improving to 87\% in under
  10 seconds. In both datasets we detect actual type errors about 82\%
  of the time, unbound variables or constructors 3--4\% of the time, and
  diverging loops 2--3\% of the time. For the remaining 11--12\% of the
  programs we are unable to provide any useful feedback.  }
\label{fig:results-witness}
\end{figure*}

\paragraph{Results}
\label{sec:results-witness}
The results of our experiments are summarized in
Figure~\ref{fig:results-witness}.
%
In both datasets our tool was able to find a witness for 83\% of the
programs in under one second, \ie\ fast enough to be integrated as a
compile-time check. If we extend our tolerance to a 10 second timeout,
we hit a maximum of 87\% coverage.
%
Interestingly, while the vast majority of witnesses corresponded to a
type-error, as expected, 3--4\% triggered an unbound variable error (even
though \ocaml\ reported a type error) and 2--3\% triggered an infinite
recursion error.
%
For the remaining 12\% of programs we were unable to provide any useful
feedback as they either completed 1,000 tests successfully, or timed out
after one minute.
%
% XX programs were deemed safe and XX timed out even at 3,000 steps, \ie
% we could not provide any useful feedback for XX\% of the total programs.
%
While a more advanced search procedure, \eg\ dynamic-symbolic execution,
could likely trigger more of the type errors, our experiments suggest that
type errors are coarse enough (or that novice programs are \emph{simple}
enough) that these techniques are not necessary.

@@

\subsection{Measuring Witness Utility}
\label{sec:user-study}
Finally, to test the explanatory power of our jump-compressed traces, we
ran a user study (IRB \#XXXX) at the University of Virginia (UVA).
%
We included four problems in an exam in the Spring session of UVA's
undergraduate Programming Languages course (CS XXX).
%
We presented the students with ill-typed \ocaml\ programs and
asked them to
%
(1) \emph{explain} the type error, and
%
(2) \emph{fix} the type error.
%
For each problem the student was given the ill-typed program and
either \ocaml's error message or \toolname's jump-compressed trace.
%
The four problems assigned to the students were the @sumList@,
@digitsOfInt@, and @wwhile@ programs from \S~\ref{sec:advantage-traces},
as well as the following @append@ program
%
\begin{ecode}
  let append x l =
    match x with
    | []   -> l
    | h::t -> h :: t :: l
\end{ecode}
%
which triggers an occurs-check error on line 4.
%
We then instructed four annotators (one of whom is an author, the other
three are teaching assistants at UCSD) to classify the answers as
correct or incorrect.

@@

\paragraph{Threats to Validity}
\subparagraph{Construct}
%
Measuring understanding is a difficult task; we used the correctness of
the student's explanation of, and fix for, the type error as a proxy for
her understanding, but it is possible that other metrics would produce
different results.

\subparagraph{Internal}
%
We assigned students randomly to two groups. The first group was given
\ocaml's errors for @append@ and \hbox{@digitsOfInt@,} and \toolname's trace
for @sumList@ and @wwhile@. The second group was given the opposite
assignment of errors and traces. This assignment ensured that (1) each
student was given \ocaml and \toolname problems, and (2) each student
was given an ``easy'' and ``hard'' problem for both \ocaml and
\toolname. Students without sufficient knowledge of \ocaml could affect
the results, as could the time-constrained nature of an exam. For these
reasons we excluded any answers left blank from our analysis.

\subparagraph{External}
%
Our experiment is based on students in the process of learning \ocaml,
and thus may not generalize to all developers. The four
programs we used were chosen manually, via a random selection and
filtering of the programs in the \ucsdbench dataset. In some cases we made
minor simplifying edits (\eg alpha-renaming, dead-code removal) to the
programs to make them more understandable in the short timeframe of an
exam; however, we never altered the resulting type-error. A different
selection of programs may lead to different results.

\subparagraph{Conclusion}
%
We collected exams from 60 students, though due to the nature of the
study not every student completed every problem. The number of
incomplete submissions ranges from 2 (for the \ocaml version of
@sumList@) to 17 (for the \toolname version of @wwhile@). Collecting
more responses per test pair was not possible, as it would require
having students answer the same problem twice (once with \ocaml and once
with \toolname).

\subsubsection{Results}
%
Figure~\ref{fig:results-user-study} summarizes a single annotator's
results, which show that students given \toolname's jump-compressed
trace were consistently more likely to correctly explain and fix the
type error than those given \ocaml's error message.
%
The measured kappa values were $\kappa = 0.72$ for the explanations and
$\kappa = 0.83$ for the fixes; while there is no formal notion for what
consititutes strong agreement~\cite{Krippendorff2012-wd}, kappa values
above $0.60$ are often called ``substantial''
agreement~\cite{Landis1977-ey}.
%
\begin{figure*}[ht]
\centering
\includegraphics[width=0.49\linewidth]{user-study-reason.png}
\includegraphics[width=0.49\linewidth]{user-study-fix.png}
\caption{A classification of students' explanations and fixes for type
  errors, given either \ocaml's error % message
  or \toolname's
  jump-compressed trace. The students given \toolname's jump-compressed
  trace consistently scored better than those given \ocaml's type
  error.}
\label{fig:results-user-study}
\end{figure*}
%
